Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube

### 2022-09-13 latex/Chapters/DRLSearch.tex
last cleanup

\newlength\mylen
\settowidth\mylen{DeepLearningHeuristic(}
\teal
\begin{algorithm}[H]
\caption{Deep Learning -- Heuristic}\label{alg:TheoryDLDRL}
\begin{algorithmic}
\Function{DeepLearningHeuristic}{**args}
\State{$puzzle\_type \gets \Call{FromParams}{**args}$}
\State{$n\_puzzles \gets$ \Call{FromParams}{**args}}
\State{$H \gets \Call{FromParams}{**args}$} \teal{\algorithmiccomment{H -- teacher heuristic to extrapolate via \textbf{DL}}}\black
\State{$loss\_function \gets \Call{FromParams}{**args}$}
\State{$net\_architecture \gets \Call{FromParams}{**args}$}
\State $V = \{state: \Call{H}{state} \ for \ state \in \Call{GenerateRandomStates}{puzzle\_type, n\_puzzles}\}$
\ \ \ \ \Return{$\Call{TrainNeuralNetwork}{V, loss\_function, net\_architecture}$}
\EndFunction
\end{algorithmic}
\end{algorithm}
\black
\noindent An alternative approach to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}). The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states. This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}. The subtlety is that both the left-hand-side and right-hand-side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking. For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops. During iteration over the first (inner) loop, the right hand-side $V$ of the value-function update equation is computed using a fixed \textit{target-network}. The left-hand-side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached. When convergence in the inner loop is deemed obtained, we break out of it, and run the outer loop, which updates the \textit{target-network} via a copy of the \textit{current-network}. The outer loop is itself broken out of when deemed appropriate (some kind of convergence criteria). See pseudo-code \ref{alg:TheoryDRL} below:
\\
\\
\teal
\begin{algorithm}[H]
\caption{Deep Reinforcement Learning -- Heuristic}\label{alg:TheoryDRL}
\begin{algorithmic}
\Function{DeepReinforcementLearningHeuristic}{**args}
\State{$puzzle\_type \gets$ \Call{FromParams}{**args}}
\State{$ puzzles\_gen\_params \gets$ \Call{FromParams}{**args}}
\State{$puzzles\_generation\_criteria \gets$ \Call{FromParams}{**args}}
\State{$n\_puzzles \gets \gets$ \Call{FromParams}{**args}}
\State{$target\_network\_update\_params \gets$ \Call{FromParams}{**args}}
\State{$max\_epochs \gets$ \Call{FromParams}{**args}}
\State{$loss\_function \gets$ \Call{FromParams}{**args}}
\State{$net\_architecture \gets$ \Call{FromParams}{**args}}
\\ \teal{\algorithmiccomment{initial network and target network}}\black
\State {$net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
\State {$target\_net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
\State{$epoch \gets 0$}
\State{$puzzles \gets null$}

\While{epoch < max\_epochs \&\& other\_convergence\_criteria(network)}
        \State{$epoch \gets epoch + 1$}
\\ \teal{\algorithmiccomment{generate \textit{training} puzzles if criteria met}}\black
        \If{puzzles\_generation\_criteria($puzzles$)}
            \State{$puzzles \gets generate\_random\_states(puzzle\_type)$}
        \EndIf
        \State{$V \gets dict()$}
        \For{$puzzle \in puzzles}$
            \State{$V[puzzle] \gets \infty$}
            \If \Call{IsGoal}{$puzzle$}:
                \State{$V[puzzle] \gets 0$}
            \Else:
	     \For{$transition, child \in \Call{Children}{$puzzle$}$}
\\ \teal{\algorithmiccomment{perform value iteration}}\black
		\State{$child_cost \gets \Call{Cost}{transition} + target\_net(child)$}
                	\State{$V[puzzle] \gets \Call{Min}{V[puzzle], child\_cost}$}
	     \EndFor
	\EndIf
       \EndFor
\\ \teal{\algorithmiccomment{iterate through network training running feed-forward and backward-prop}}\black
	\State{$net \gets \Call{TrainNeuralNet}{V, loss\_function, net\_architecture}$}
\\ \teal{\algorithmiccomment{update target network if criteria met}}\black
	\If{$target\_network\_update\_criteria(net, target\_net)$}
		\State{$target\_net \gets \Call{Copy}{net}$}
	\EndIf
    \Return{$net$}
\EndWhile

@@

\\ \teal{\algorithmiccomment{ ... same params and init as for \textbf{DRL} ...}}\black
 \State{$nb\_actions \gets$ \Call{GetNbActions}{puzzle\_type}}
\While{epoch ...}
\\ \teal{\algorithmiccomment{ ... same while epoch loop as for \textbf{DRL} ...}}\black
\\ \teal{\algorithmiccomment{ ... update cost \textbf{and actions} via V \& Q iteration ...}}\black
	\State{$Q\_and\_V \gets dict()$}
	\For {$puzzle \in puzzles$}
            \State{$value \gets target\_network(puzzle).value$}
            \State{$actions = [0] * nb\_actions$}
            \State{$best\_action\_id \gets 0$}
	 \If{\Call{IsGoal}{$puzzle$}}
                \State{$value \gets 0$}
            \Else:
                \For{$action\_id, child \in \Call{Children}{puzzle}$}
                    \State{$child\_value \gets target\_network(child).value$}
                    \If{$child\_value \le value$}:
                        \State{$value \gets child\_value$}
                        \State{$best\_action\_id \gets action\_id$}
		\EndIf
		\EndFor
	\EndIf
\\ \teal{\algorithmiccomment{ ... update cost \textbf{and best action} ...}}\black
            \State{$actions[best\_action\_id] \gets 1$}
            \State{$Q\_and\_V[puzzle] \gets actions + [value]$}
	\EndFor
\EndWhile
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

@@

The \textbf{DQL} heuristics from \ref{alg:TheoryDQL} no longer only produce cost-to-go estimates as \textbf{DRL}, but rather a joint cost-to-go and distribution over actions (i.e. a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm combining the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. It generates at each iteration (possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (allowing exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$. To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and down-weight their probability so that subsequent paths explore different actions. Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
\\

@@

\teal
\begin{algorithm}[H]
\caption{Monte Carlo Tree Search}\label{alg:TheoryMCTS}
\begin{algorithmic}
\Function{MCTS}{$root, time\_out, q\_v\_network, c$}
\State{\Call{InitialiseTimeOut}{$time\_out$}}
\State{$tree \gets \Call{Tree}{root}$}
\If{\Call{IsGoal}{$root$}}
  \Return{$root$}
\EndIf
\While{true}
	\State\Call{CheckTimeOut}{void}
	\State{$leaf \gets root$}
	\While{$leaf.expanded$}
        		\State{$actions\_probas, actions\_values \gets q\_v\_network(leaf)$}
	           \State{$leaf \gets\Call{ChooseChildNode}{leaf,  actions\_probas, actions\_values, leaf.actions\_count, c}$}
		 \State{\Call{Expand}{$leaf, tree$}}
	\EndWhile
           \Call{IncrementActionsFromLeafToRoot}{$leaf$}
       \\ \teal{\algorithmiccomment{path to leaf can usually be improved by re-running \textbf{BFS} on the resulting tree}}\black
       \If{\Call{IsGoal}{$leaf$}}
            \Return{\Call{BlindSearch}{$tree, search\_type=Search.BFS$}}
       \EndIf
\EndWhile
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

\noindent Note that for this project, I have implemented all the above algorithms in Python (BFS, DFS, A$^{*}$, DeepLearner, DeepReinforcementLearner, DeepQLearner, MonteCarloTreeSearch), and more details can be found in section (\ref{sec:Implementation}).

>>>>

\newlength\mylen
\settowidth\mylen{DeepLearningHeuristic(}
\teal
\begin{algorithm}[H]
\caption{Deep Learning -- Heuristic}\label{alg:TheoryDLDRL}
\begin{algorithmic}
\Function{DeepLearningHeuristic}{**args}
\State{$puzzle\_type \gets \Call{FromParams}{**args}$}
\State{$n\_puzzles \gets$ \Call{FromParams}{**args}}
\State{$H \gets \Call{FromParams}{**args}$} \teal{\algorithmiccomment{H -- teacher heuristic to extrapolate via \textbf{DL}}}\black
\State{$loss\_function \gets \Call{FromParams}{**args}$}
\State{$net\_architecture \gets \Call{FromParams}{**args}$}
\State $V = \{state: \Call{H}{state} \ for \ state \in \Call{GenerateRandomStates}{puzzle\_type, n\_puzzles}\}$
\ \ \ \ \Return{$\Call{TrainNeuralNetwork}{V, loss\_function, net\_architecture}$}
\EndFunction
\end{algorithmic}
\end{algorithm}
\black
\noindent An alternative approach to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}). The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states. This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}. The subtlety is that both the left-hand-side and right-hand-side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking. For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops. During iteration over the first (inner) loop, the right hand-side $V$ of the value-function update equation is computed using a fixed \textit{target-network}. The left-hand-side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached. When convergence in the inner loop is deemed obtained, we break out of it, and run the outer loop, which updates the \textit{target-network} via a copy of the \textit{current-network}. The outer loop is itself broken out of when deemed appropriate (some kind of convergence criteria). See pseudo-code \ref{alg:TheoryDRL} below:
\\
\\
\teal
\begin{algorithm}[H]
\caption{Deep Reinforcement Learning -- Heuristic}\label{alg:TheoryDRL}
\begin{algorithmic}
\Function{DeepReinforcementLearningHeuristic}{**args}
\State{$puzzle\_type \gets$ \Call{FromParams}{**args}}
\State{$ puzzles\_gen\_params \gets$ \Call{FromParams}{**args}}
\State{$puzzles\_generation\_criteria \gets$ \Call{FromParams}{**args}}
\State{$n\_puzzles \gets $ \Call{FromParams}{**args}}
\State{$target\_network\_update\_params \gets$ \Call{FromParams}{**args}}
\State{$max\_epochs \gets$ \Call{FromParams}{**args}}
\State{$loss\_function \gets$ \Call{FromParams}{**args}}
\State{$net\_architecture \gets$ \Call{FromParams}{**args}}
\\ \teal{\algorithmiccomment{initial network and target network}}\black
\State {$net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
\State {$target\_net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
\State{$epoch \gets 0$}
\State{$puzzles \gets null$}

\While{epoch < max\_epochs \&\& other\_convergence\_criteria(net)}
        \State{$epoch \gets epoch + 1$}
\\ \teal{\algorithmiccomment{generate \textit{training} puzzles if criteria met}}\black
        \If{puzzles\_generation\_criteria($puzzles$)}
            \State{$puzzles \gets generate\_random\_states(puzzle\_type, n\_puzzles)$}
        \EndIf
        \State{$V \gets dict()$}
        \For{$puzzle \in puzzles}$
            \State{$V[puzzle] \gets \infty$}
            \If{\Call{IsGoal}{$puzzle$}}:
                \State{$V[puzzle] \gets 0$}
            \Else:
	     \For{$transition, child \in \Call{Children}{$puzzle$}$}
\\ \teal{\algorithmiccomment{perform value iteration}}\black
		\State{$child\_cost \gets \Call{Cost}{transition} + target\_net(child)$}
                	\State{$V[puzzle] \gets \Call{Min}{V[puzzle], child\_cost}$}
	     \EndFor
	\EndIf
       \EndFor
\\ \teal{\algorithmiccomment{iterate through network training running feed-forward and backward-prop}}\black
	\State{$net \gets \Call{TrainNeuralNet}{V, loss\_function, net}$}
\\ \teal{\algorithmiccomment{update target network if criteria met}}\black
	\If{$target\_network\_update\_criteria(net, target\_net)$}
		\State{$target\_net \gets \Call{Copy}{net}$}
	\EndIf
\EndWhile
\Return{$net$}

@@

\\ \teal{\algorithmiccomment{ ... same params and init as for \textbf{DRL} ...}}\black
 \State{$nb\_actions \gets$ \Call{GetNbActions}{puzzle\_type}}
\While{epoch ...}
\\ \teal{\algorithmiccomment{ ... same while epoch loop as for \textbf{DRL} ...}}\black
\\ \teal{\algorithmiccomment{ ... update cost \textbf{and actions} via V \& Q iteration ...}}\black
	\State{$Q\_and\_V \gets dict()$}
	\For {$puzzle \in puzzles$}
            \State{$value \gets target\_network(puzzle).value$}
            \State{$actions = [0] * nb\_actions$}
            \State{$best\_action\_id \gets 0$}
	 \If{\Call{IsGoal}{$puzzle$}}
                \State{$value \gets 0$}
            \Else:
                \For{$action\_id, child \in \Call{Children}{puzzle}$}
                    \State{$child\_value \gets target\_network(child).value$}
                    \If{$child\_value \le value$}:
                        \State{$value \gets child\_value$}
                        \State{$best\_action\_id \gets action\_id$}
		\EndIf
		\EndFor
	\EndIf
\\ \teal{\algorithmiccomment{ ... update cost \textbf{and best action} ...}}\black
            \State{$actions[best\_action\_id] \gets 1$}
            \State{$Q\_and\_V[puzzle] \gets actions + [value]$}
	\EndFor
\\ \teal{\algorithmiccomment{ ... same as \textbf{DRL} ...}}\black
\State{$net \gets \Call{TrainNeuralNet}{Q\_and\_V, loss\_function, net}$}
\EndWhile
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

@@

The \textbf{DQL} heuristics from \ref{alg:TheoryDQL} no longer only produce cost-to-go estimates as \textbf{DRL}, but rather a joint cost-to-go and distribution over actions (i.e. a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm, such as the Monte Carlo Tree Search (\textbf{MCTS}), combining the actions distribution and the value function to inform its search. \textbf{MCTS} generates at each iteration (possibly in a distribute manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until it is). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$. To prevent always following the same paths, the \textbf{MCTS} keeps track of actions already followed and down-weights their probability so that subsequent paths explore different actions. Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
\\

@@

\teal
\begin{algorithm}[H]
\caption{Monte Carlo Tree Search}\label{alg:TheoryMCTS}
\begin{algorithmic}
\Function{MCTS}{$root, time\_out, q\_v\_network, c$}
\State{\Call{InitialiseTimeOut}{$time\_out$}}
\If{\Call{IsGoal}{$root$}}
\State{$tree \gets \Call{Tree}{root}$}
  \Return{$root$}
\EndIf
\While{true}
	\State\Call{CheckTimeOut}{void}
	\State{$leaf \gets tree.root$}
	\While{$leaf.expanded$}
        		\State{$actions\_probas, actions\_values \gets q\_v\_network(leaf)$}
	           \State{$leaf \gets\Call{ChooseChildNode}{leaf,  actions\_probas, actions\_values, leaf.actions\_count, c}$}
		 \State{\Call{Expand}{$leaf, tree$}}
	\EndWhile
           \Call{IncrementActionsFromLeafToRoot}{$leaf$}
       \\ \teal{\algorithmiccomment{path to leaf can usually be improved by re-running \textbf{BFS} on the resulting tree}}\black
       \If{\Call{IsGoal}{$leaf$}}
            \Return{\Call{BlindSearch}{$tree, search\_type=Search.BFS$}}
       \EndIf
\EndWhile
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

\noindent Note that for this project, I have implemented all the above algorithms in Python (\textbf{BFS}, \textbf{DFS}, A$^{*}$, DeepLearner, DeepReinforcementLearner, DeepQLearner, MonteCarloTreeSearch), and more details can be found in section \ref{sec:Implementation}.

### 2022-09-13 latex/Chapters/Implementation.tex
last cleanup

\Section{rubiks.search}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesearch}
%\decoRule
\caption[Codebase]{rubiks.search}
\label{fig:Codebasesearch}
\end{figure}
This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
\begin{itemize}
\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
\end{itemize}

@@

\Section{rubiks.solvers}
\label{sec:codesolvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements the actual puzzles solvers. The base class Solver is a Factory, and in addition to being able to instantiate the following types of solvers, can run different solvers through a similar sequence of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver). Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible, so I used A$^{*}$ with KociembaHeuristic as my benchmark. For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{alg:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{alg:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{alg:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} My implementation follows McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470} but is however single-threaded. As a consequence, I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths (in single threaded mode, this parameter is inconsequential).
\\
This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process. It is indeed clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops. I have however made this \textbf{BFS} trimming optional via a \textit{trim\_tree} parameter to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
\\
I will study and report on the effect of the trade-off hyper-parameter \textit{c}, and confirm how it affects solutions' quality and run time in section \ref{sec:MTCSHyperParamsEffect}.

>>>>

\Section{rubiks.search}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesearch}
%\decoRule
\caption[Codebase]{rubiks.search}
\label{fig:Codebasesearch}
\end{figure}
This modules contains graph search strategies. I have actually reused some of the code I implemented for an \textbf{AIPnT} assignment here. It contains the following classes:
\begin{itemize}
\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
\end{itemize}

@@

\Section{rubiks.solvers}
\label{sec:codesolvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.22]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements the actual puzzles solvers. The base class Solver is a Factory, and in addition to being able to instantiate the following types of solvers, can run different solvers through a similar sequence of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver). Ideally the benchmark solver to compute the optimality score should be provably optimal, but in the case of the \textbf{RC} that was not possible, so I used Kociemba as benchmark. For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{alg:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{alg:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{alg:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} My implementation follows McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470} but is however single-threaded. As a consequence, I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths (in single threaded mode, this parameter is inconsequential).
\\
This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf. Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process. It is indeed clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops. I have however made this \textbf{BFS} trimming optional via a \textit{trim\_tree} parameter to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
\\
I will study and report on the effect of the trade-off hyper-parameter \textit{c}, and confirm how it affects solutions' quality and run time in section \ref{sec:MTCSHyperParamsEffect}.
